On using the decision trees to identify the local extrema in parallel global optimization algorithms \footnote{This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .}

### 2023-03-20 2023/PAVTShort/ParallelDecisionTreeShort.tex
Короткая версия статьи по деревьям

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
\usepackage{marvosym}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\usepackage[english]{babel} 
\usepackage{amsmath} 
\def\UrlFont{\rmfamily}
\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
%
\titlerunning{On using the decision trees to identify the local extrema }  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and Dmitry Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
%
\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{K.A.~Barkalov and I.G.~Lebedev, and D.I.~Silenko}
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
\email{ \{konstantin.barkalov,ilya.lebedev,silenko\}@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered. The objective function of the  problem is defined as a ``black box''.  We assume it only to satisfy Lipschitz condition with unknown constant.  Global search algorithm is applied for the search of global minimum in the problems of such type. In the present work, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information. Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.  This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Machine learning $\cdot$ Decision tree}
\end{abstract}
%

@@

	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute the value $R(i)$ called \textit{a  characteristic} of the interval according to the formulae 
	\begin{equation} 
		\label{agp3_R1} R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}, 
	\end{equation} 
	\begin{equation} 
		\label{agp3_Ri} R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1, 
	\end{equation} 
	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

@@

\begin{figure}[ht!]

	\begin{center} 
				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.pdf}
					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 	
					 			\label{fig:fig3}
				 \end{minipage}
	\end{center}
\end{figure}

@@

As one can see from Table \ref{tab:2},  the iteration speedup is large enough for the problems of any  dimensionality. 
So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.  Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree. The use of such a combination of methods allows considerably accelerating the algorithm.

@@

\medskip

\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .

>>>>

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{pavt-en} 
\usepackage{amsmath}

\begin{document}

% \title - название статьи
% \authors - список авторов

\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms  \footnote{This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .}}

\authors{Konstantin Barkalov\superscript{1},Ilya Lebedev\superscript{1}, Dmitry Silenko\superscript{1}}
\organizations{\superscript{1}Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}

% Аннотация заключается в окружение abstract
\begin{abstract}
In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered. The objective function of the  problem is defined as a ``black box''.  We assume it only to satisfy Lipschitz condition with unknown constant.  Global search algorithm is applied for the search of global minimum in the problems of such type. In the present work, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information. Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.  This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
\end{abstract}

\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Machine learning $\cdot$ Decision tree}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня
% Не используйте уровень вложенности заголовков больше трех!
% Каждый абзац текста в статье начинается командой \par или пустой
% строкой.

@@

	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute the value $R(i)$ called \textit{a  characteristic} of the interval according to the formulae

	\begin{equation} 
		\label{agp3_R1} R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}, 
	\end{equation}

	\begin{equation} 
		\label{agp3_Ri} R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1, 
	\end{equation}

	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

@@

@@

As one can see from Table \ref{tab:2},  the iteration speedup is large enough for the problems of any  dimensionality.

\section{Conclusion}

So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.  Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree. The use of such a combination of methods allows considerably accelerating the algorithm.

@@
